6.867 Problem Set 1

### 2015-10-02 pset1/paper.tex
wrote 1, part of 2

>>>>

\section{Linear Basis Function Regression}
Given a feature matrix $\Phi$, a weight vector $\theta$, and an output vector $Y$, we define the sum-of-squares error (SSE) of $\theta$ as
$$(\Phi\theta - Y)^T(\Phi\theta - Y).$$
The particular vector $\theta$ that minimizes this value is
$$\theta = (\Phi^T\Phi)^{-1}\Phi^TY.$$
The gradient of the sum-of-squares error with respect to $\theta$ is
$$2\Phi^T(\Phi\theta - Y).$$
We implemented functions to calculate, using the closed forms given above, both the SSE-minimizing value (i.e. the value with the maximum likelihood) of $\theta$ as well as the gradient of the SSE with respect to $\theta$. We verified the correctness of the maximum-likelihood function by checking it against data sets for which the maximum likelihood vectors were known. We also verified the correctness of the gradient by checking it against a numerical implementation.
